---
title: "Regular expressions"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE)
options(dplyr.summarise.inform = FALSE)
```

```{r load_packages, message=FALSE, warning=FALSE}
library(tidyverse)
```
This notebook contains some examples on the use of regular expressions and the tidyverse string functions contained in`stringr`
It is by no means comprehensive - regular expressions can be considered a language in their own right. See:
https://stringr.tidyverse.org/articles/regular-expressions.html
```
======
For `stringr` see:
https://stringr.tidyverse.org/articles/stringr.html
------------------------------------------------------------------
## Detecting
```{r}
test_str <- c("Paul",
              "paul",
              "SY23 5ND",
              "SY235ND",
              "25/12/2019",
              "25/12/19")
```
------------------------------------------------------------------
```{r}
# Match "Paul"

pattern <- "Paul"

test_str %>% 
  str_detect(pattern)
```
------------------------------------------------------------------
```{r}
# Match "Paul or paul"

pattern <- "(?i)Paul"

test_str %>% 
  str_detect(pattern)
```
------------------------------------------------------------------
```{r}
# Match postcode with a space.

pattern <- "\\w{1,2}\\d{1,2} \\d{1,2}\\w{1,2}"

test_str %>% 
  str_detect(pattern)
```
-------------------------------------------------------------------
```{r}
# Match a date in the format dd/mm/yyyy

pattern <- "\\d{1,2}/\\d{1,2}/\\d{4}"

test_str %>% 
  str_detect(pattern)
```
-------------------------------------------------------------------
```{r}
# Match a date in the format dd/mm/yyyy or dd/mm/yy

pattern <- "\\d{1,2}/\\d{1,2}/(\\d{4}|\\d{2})"

test_str %>% 
  str_detect(pattern)
```
-------------------------------------------------------------------

```{r}
# Extract the first date from sentence.

sentence <- "The holiday will start on 24/12/2019 and end on 02/01/2020"

pattern <- "\\d{1,2}/\\d{1,2}/(\\d{4}|\\d{2})"

sentence %>% 
  str_extract(pattern)
```
-------------------------------------------------------------------




























