---
title: "Regular expressions"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE)
options(dplyr.summarise.inform = FALSE)
```

```{r load_packages, message=FALSE, warning=FALSE}
library(tidyverse)
```

This notebook contains some examples on the use of regular expressions and the tidyverse string functions contained in`stringr`

It is by no means comprehensive - regular expressions can be considered a language in their own right. See:

https://stringr.tidyverse.org/articles/regular-expressions.html


For `stringr` see:

https://stringr.tidyverse.org/articles/stringr.html

# Examples

## Detecting

```{r}
test_str <- c("Paul",
              "paul",
              "SY23 5ND",
              "SY235ND",
              "25/12/2019",
              "25/12/19")
```

```{r}
# Match "Paul"

pattern <- "Paul"

test_str %>% 
  str_detect(pattern)
```


```{r}
# Match "Paul or paul"

pattern <- "(?i)Paul"

test_str %>% 
  str_detect(pattern)
```

```{r}
# Match postcode with a space.

pattern <- "\\w{1,2}\\d{1,2} \\d{1,2}\\w{1,2}"

test_str %>% 
  str_detect(pattern)
```

```{r}
# Match postcode with or without a space.
 
pattern <- "\\w{1,2}\\d{1,2} ?\\d{1,2}\\w{1,2}"

test_str %>% 
  str_detect(pattern)
```

```{r}
# Match a date in the format dd/mm/yyyy

pattern <- "\\d{1,2}/\\d{1,2}/\\d{4}"

test_str %>% 
  str_detect(pattern)
```

```{r}
# Match a date in the format dd/mm/yyyy or dd/mm/yy

pattern <- "\\d{1,2}/\\d{1,2}/(\\d{4}|\\d{2})"

test_str %>% 
  str_detect(pattern)
```

## Extracting

```{r}
# Extract the first date from sentence.

sentence <- "The holiday will start on 24/12/2019 and end on 02/01/2020"

pattern <- "\\d{1,2}/\\d{1,2}/(\\d{4}|\\d{2})"

sentence %>% 
  str_extract(pattern)
```

```{r}
# Extract both dates from sentence.

sentence %>% 
  str_extract_all(pattern)
```

```{r}
# Extract the day month and year individually using capture groups

pattern <- pattern <- "(\\d{1,2})/(\\d{1,2})/(\\d{4}|\\d{2})"

dd_mm_yyyy <- 
sentence %>% 
  str_match(pattern)

dd_mm_yyyy

dd_mm_yyyy[,2]
dd_mm_yyyy[,3]
dd_mm_yyyy[,4]
```


Actually this isn't a good example, because we have `lubridate` to handle dates.

```{r, hold = TRUE}
library(lubridate, warn.conflicts = FALSE)
one_more_sleep <- dd_mm_yyyy[,1]
one_more_sleep
class(one_more_sleep)
one_more_sleep <- dmy(one_more_sleep)
one_more_sleep
class(one_more_sleep)
day(one_more_sleep)
wday(one_more_sleep, label = TRUE)
month(one_more_sleep)
year(one_more_sleep)
class(year(one_more_sleep))
typeof(year(one_more_sleep))
```

Extraction from brackets

```{r}
str_match("The host country [England].", "\\[(\\w+)\\]")
```

