---
title: "Cleaning"
---

## Some examples on cleaning

```{r}
library(tidyverse)
library(lubridate)
```

## Rownames

If we look at mtcars ther is one obvious omission - the make and model of the car.

```{r}
mtcars %>% 
  tbl_df()
```

This is because the make and model is stored in the row names - a bad idea.

```{r}
mtcars %>% 
  row.names()
```

We can correct this:

```{r}
(mtcars_proc <- 
  mtcars %>%
  rownames_to_column("Model"))
```

We might want to separate make and model.

```{r}
(mtcars_proc <- 
   mtcars_proc %>% 
   mutate(Model = str_trim(Model)) %>% 
   separate(Model, into = c("Make", "Model"), sep = "\\s", extra = "merge"))
```

## A quick example on dates and the use of `lubridate`


Create some dummy data

```{r}
set.seed(123)
(dat_raw <- 
 tibble(Dy = sample(1:31, 50, replace = TRUE,),
          Mnth = sample(1:12, 50, replace = TRUE,),
          Yr = sample(1960:2019, 50, replace = TRUE,)) %>% 
   mutate(Date = paste(Dy, Mnth, Yr, sep = ".")))

```

## Convert to a date type.

```{r}
(dat_proc <- 
   dat_raw %>% 
   mutate(Date = dmy(Date)))
```
(Why did one fail to parse?)

Once we have a data format we can use to lubridate functions to extract day, month and year.

```{r}
day(dat_proc$Date)
```

```{r}
wday(dat_proc$Date, label = TRUE)
```

```{r}
month(dat_proc$Date, label = TRUE)
```









